Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-01-09 komlos.tex

%Here are some examples:
%\begin{itemize}
%\item in approximate counting we have $f(x, q) = 1$ if $x=q$ and zero else. Here both $x$ and $q$ belong to some finite domain.
%\item in quantile approximation we have $f(x, q) = 1$ if $x<q$ and zero else. Here, $x$ and $q$ belong to a set which exhibits total ordering.
%\item in matrix column subset selection we have $f(x, q) = \langle x,q \rangle ^2$ where $q$ is any unit vector in $\R^d$ and $x$ are matrix rows. For simplicity, we will assume in this manuscript that $x$ is also unit norm.
%\item in density estimation $f(x, q) = \exp(- \|x-y\|_2^2/\sigma^2)$ or any other kernel (see \cite{})
%\item in logistic regression $f(x, q) = \operatorname{loss}(x,q) = \log(1 + \exp(-q^T x))$. 
%Here $q$ is the parameters of the linear model and $x$ is a training example vector multiplied by its label ($1$ or $-1$).
%\item Linear Classification $f(x,q) = 1$ if $x^Tq > 0$ and $0$ else. Here, like in Logistic Regression, $x$ is the example point multiplied by its label and $q$ is the weights of the linear classifier.
%\item Fractional Satisfiability $f(x,q)$ if $q$ satisfies $f$ and $0$ otherwise. Here $x$ in a clause in conjunctive normal form and $q$ is a boolean value assignment to the variables. For example $f(x,q) = 1$ iff $(q_1 \vee q_5  \vee \neg q_8)$ is true.
%\item Quadratic forms of the graph Laplacian are $f(x, q) = q^T L_x q = \sum_{i,j\in E} (q_i - q_i)^2$ where $x$ is an edge and $L_x$ is the graph laplacian corresponding to the graph with a single edge $x$ and $q$ is a test vector.
%\item Probably also good for matrix approximation when we get entry updates. 
%\end{itemize}

For bounded functions $f$ like above, uniform sampling of $\ell = O(\log(1/\delta)/\eps^2)$ combined with a union bound over $|\mathcal Q|$ always provides a valid solution using $\ell = O(\log(|\mathcal Q|)/\eps^2)$ stream items. 
While $|\mathcal Q|$ is often infinite it reduces to a finite (albeit usually exponentially large) set through standard epsilon net arguments. 
%This is common practice in machine learning, specifically, PAC learning. However, it is often far from being the smallest cardinality coreset possible.
We argue that we can offer a solution for producing smaller coresets for all of the above problems in a unified manner. 
Moreover, our solution creates streaming algorithms with fully mergeable sketches. 
The size of the coreset above appears to be intimately tied to the discrepancy properties of the associated functions.
This is captured by its Komlos complexity.

@@

\section{Komlos Complexity}
We begin by formally defining the Komols complexity and by giving some general results about for problems with low Komols complexity. The Komols complexity of a sequence set is defined similarly to it Rademacher complexity.
%\begin{definition}
%Let $A \subset [0,1]^m$ and $\sigma \in \{-1,1\}^m$ uniformly. 
%The Rademacher complexity of $\F$ is 
%\[
%R_m(\F) =  \frac{1}{m}\E_{\sigma} \sup_{a \in A}  \sum_{i=1}^{m}\sigma_i a_i
%\]
%\end{definition}
%
%\begin{definition}
%Let $A \subset [0,1]^m$ and $\sigma \in \{-1,1\}^m$. 
%The Komlos complexity of $A$ is 
%\[
%K_m(A) =  \frac{1}{m}\min_{\sigma} \sup_{a \in A}  \left|\sum_{i=1}^{m}\sigma_i a_i\right|
%\]
%\end{definition}

>>>>

For bounded functions $f$ like above, uniform sampling of $\ell = O(\log(1/\delta)/\eps^2)$ combined with a union bound over $|\mathcal Q|$ always provides a valid solution using $\ell = O(\log(|\mathcal Q|)/\eps^2)$ items. 
While $|\mathcal Q|$ is often infinite it reduces to a finite (albeit usually exponentially large) set through standard epsilon net arguments. 
%
We argue that we can offer a solution for producing smaller coresets for all of the above problems in a unified manner. 
Moreover, our solution creates streaming algorithms with fully mergeable sketches. 
The size of the coreset above appears to be intimately tied to the discrepancy properties of the associated functions.
This is captured by its Komlos complexity.

@@

\section{Komlos Complexity}
We begin by formally defining the Komols complexity and by giving some general results about for problems with low Komols complexity. The Komols complexity of a sequence set is defined similarly to it Rademacher complexity.
